No Correlations were Found between the efforts made by the Staff of Toronto's Central Intake Line and Toronto's Homeless Death Counts

### 2024-01-23 outputs/paper/paper.qmd
finished the writing portion of the paper

---
title: "No Correlations were Found between the efforts made by the Staff of Toronto's Central Intake Line and Toronto's Homeless Death Counts"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/central_intake_calls_and_homeless_death_count.git"
date: today
date-format: long
abstract: "For the City of Toronto, the physical well-being of its local homeless population has always been one of its primary focuses. Numerous social efforts, funded by the City, were made in attempts to reduce the death rate of the local homeless population. In order to visualize the magnitude of social impact enforced by the City of Toronto, analyses were performed to investigate the correlation between the efforts made by the staff of Toronto’s Central Intake Line and the monthly homeless death counts between November 2020 to June 2023 using data from OpenDataToronto. Staff efforts such as the number of calls coded, number of referral to shelter, and number of information for homelessness prevention, were documented and compared against the monthly homeless death counts. No discernible correlations were found between each of these efforts to the monthly homeless death counts."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

In this paper, we want to measure the magnitude of positive impact created by the Central Intake Line in relation to the monthly cumulative number of deaths in the homeless community. Using Central Intake Call Wrap-up Codes data [@Central_Intake_Call] from OpenDataToronto [@OpenDataToronto], three different aspects of the Central Intake Line efforts were isolated: total calls coded, referral to shelter, and homelessness prevention information provided. These three key aspects were measured against the monthly cumulative Deaths of People Experiencing Homelessness Data [@Homeless_Death_Count] documented by the Toronto Public Health [@TPH]. The analyses yielded no discernible correlation between each of the three efforts and the monthly death counts. Further investigation could be conducted in the future if more detailed tracking data of the individual callers became available.

Analyses and findings in this paper are structured into several sections: @sec-data -- Data, @sec-result -- Result, @sec-discussion -- Discussion, and @sec-conclusion -- Conclusion. The Data section examines all datasets and variables kept for analysis, followed by an explanation of their data cleaning processes. The Result section focuses on visualizing and presenting the correlation between desired variables. The Discussion sections further evaluate the meaning behind the correlation presented in the previous section. Lastly, the arguments presented in all previous sections are wrapped up by the Conclusion section which summarizes the main findings in this paper.

@@

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

@@

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: false
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

---
title: "No Correlations were Found between the efforts made by the Staff of Toronto's Central Intake Line and Toronto's Homeless Death Counts"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/central_intake_calls_and_homeless_death_count.git"
date: today
date-format: long
abstract: "For the City of Toronto, the physical well-being of its local homeless population has always been one of its primary focuses. Numerous social efforts, funded by the City, were made in attempts to reduce the death rate of the local homeless population. In order to visualize the magnitude of social impact enforced by the City of Toronto, analyses were performed to investigate the correlation between the efforts made by the staff of Toronto’s Central Intake Line and the monthly homeless death counts between November 2020 to June 2023 using data from OpenDataToronto. Staff efforts such as the number of calls coded, number of referral to shelter, and number of information for homelessness prevention, were documented and compared against the monthly homeless death counts. An inverse correlation is found between the number of calls coded and the homeless death count. No discernible correlations were found between coded call types of interest and the monthly homeless death counts."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

In this paper, we want to measure the magnitude of positive impact created by the Central Intake Line in relation to the monthly cumulative number of deaths in the homeless community. Using Central Intake Call Wrap-up Codes data [@Central_Intake_Call] from OpenDataToronto [@OpenDataToronto], three different aspects of the Central Intake Line efforts were isolated: total calls coded, referral to shelter, and homelessness prevention information provided. These three key aspects were measured against the monthly cumulative Deaths of People Experiencing Homelessness Data [@Homeless_Death_Count] documented by the Toronto Public Health [@TPH]. The analyses yielded an inverse correlation between the number of calls coded and the homeless death count, and no discernible correlations between coded call types of interest and the monthly homeless death counts.

Analyses and findings in this paper are structured into several sections: @sec-data -- Data, @sec-result -- Result, @sec-discussion -- Discussion, and @sec-conclusion -- Conclusion. The Data section examines all datasets and variables kept for analysis, followed by an explanation of their data cleaning processes. The Result section focuses on visualizing and presenting the correlation between variables of interest. The Discussion sections further evaluate the meaning behind the correlation presented in the previous section. Lastly, the arguments presented in all previous sections are wrapped up by the Conclusion section which summarizes the main findings in this paper.

@@

## Correlation between number of Calls Coded & Homeless Death Counts {#sec-relation-calls}
Based on @fig-coded, there seems to be an inverse relationship between the directions of change for the homeless death count trend over time, the blue line, and the monthly cumulative calls coded represented by the red lines. When the red line moves upward, the blue line moves down. This inverse relationship indicates a negative correlation between the two variables. A negative correlation between these two variables implies that the efforts the Central Intake Line placed into coding calls has created a positive impact in reducing homeless death in Toronto.

## Correlation between number of Coded Calls of Interest & Homeless Death Counts
According to the @fig-referred and @fig-informed, both the number of callers referred to shelters (represented by the purple line in @fig-referred) and the number of callers provided with information on homelessness prevention (represented by the dark green line in @fig-informed) change in the same direction as the monthly cumulative homeless death count. Through pure visual observation, a positive correlation between homeless death count to each of the variables would be found. This relationship would imply that the Central Intake Line [@Central_Intake_Call] has actually created a negative impact in reducing the homeless death count in Toronto by referring callers to shelter and providing callers with information on homelessness prevention. However, the magnitude of fluctuation for both the purple and dark green line are very large throughout the entire time period. Especially from the beginning of 2022 to the end of 2023 for the purple line. Due to the magnitude of fluctuation for both lines, we cannot responsibly conclude any causal relationship between the two variables by pure observation of the trends. We can only say that the general trend of the purple and dark green line moves in the same direction as the homeless death count trend. Therefore, no meaningful correlation between the two variables were concludable using only the Central Intake Call Wrap-Up Codes Dataset [@Central_Intake_Call] and Deaths of People Experiencing Homelessness Dataset [@Homeless_Death_Count] obtained from OpenDataToronto portal [@OpenDataToronto]. However, this result does not imply that no correlation between variables exists in the real world. It is completely plausible that correlation between the two variables can be found if more detailed data were provided, or more advanced statistical models were used for analysis.

## Errors in analysis, Limitations and Next Steps

Potential errors exist in the final datasets used for analyses. During the data cleaning process,  if rows in the Deaths of People Experiencing Homelessness Dataset [@Homeless_Death_Count] contain “Unknown” as a Month or Year value, then these rows are filtered out of the final dataset used for relationship analyses. This filtering action would cause an inaccuracy in the total cumulation of the homeless death count; however, incoherence of dates would occur if these rows were left inside the dataset. Another potential error could occur even before I download datasets from OpenDataToronto [@OpenDataToronto]. We cannot rule out the possibility of human error during the data inputting process before the data is submitted to the OpenDataToronto portal [@OpenDataToronto].

Limitations to this analysis exist for a few reasons. Firstly, each Central Intake Line staff has their own understanding about which code they should assign their calls. We have no way of verifying whether the staff have categories their calls correctly. Secondly, the staff do not have enough time to conduct repeated follow-ups with each caller they were assigned to. We have no way of verifying what percentage of callers referred to a shelter actually got there. Lastly, we do not have enough information to sort the coded calls by demography. A subset of data that sorts the homeless death count by demography exists in the Deaths of People Experiencing Homelessness Catalogue [@Homeless_Death_Count], if more information about the demography of the callers were provided, more information about the relationship between variables of interest might be found. Further analysis on homeless death counts by demography can be conducted based on shelter data that contains demographic values.

@@

This paper investigated the magnitude of impact created by the Central Intake Line in relation to the monthly cumulative number of deaths in the homeless community. Variables of interest such as: calls coded, number of callers referred to shelter, number of callers provided with information on homelessness prevention, and homeless death count, were obtained from the Central Intake Call Wrap-Up Codes Dataset [@Central_Intake_Call] and the Deaths of People Experiencing Homelessness Dataset [@Homeless_Death_Count] stored in the OpenDataToronto portal [@OpenDataToronto]. Statistical analyses yielded a negative correlation between the number of calls coded and the homeless death count, which implies a positive impact in reducing homeless death in Toronto. Other variables, such as number of callers referred to shelter and number of callers provided with information on homelessness prevention, yielded no meaningful correlation to the homeless death count. Therefore we cannot conclude the magnitude of their contribution to the reduction of homeless death count. Further analysis on homeless death counts by demography can be conducted based on shelter data that contains demographic values.
